A partisan affair: Mapping patronage in municipal bureaucracies of Brazil

### 2021-01-15 paper/draft.tex
begin outlining patronage section

There is an ongoing debate on the strength of partisan ties in Brazil. On the one hand, scholars have noted that partisanship in Brazil is weak, meaning that politicians and voters do not have strong party loyalty and often ``switch'' to other parties \citet{desposato2006parties,ames2002deadlock}. On the other, some scholars have noted that party ties have grown in strength over time, in particular for leftist programmatic parties such as the PT \citep{samuels2014power,samuels2006sources}. Part of this debate owes to disagreements on how to measure party strength, whether it be testing voters' prior knowledge of party's ideological positions or if instead, it should be measured by testing whether voters issue ballots for individual candidates or their party labels -- which disproportionately

>>>>

There is an ongoing debate on the strength of partisan ties in Brazil. On the one hand, scholars have noted that partisanship in Brazil is weak, meaning that politicians and voters do not have strong party loyalty and often ``switch'' to other parties \citet{desposato2006parties,ames2002deadlock}. On the other, some scholars have noted that party ties have grown in strength over time, in particular for leftist programmatic parties such as the PT \citep{samuels2014power,samuels2006sources}. Part of this debate owes to disagreements on how to measure party strength, whether it be testing voters' prior knowledge of party's ideological positions or if instead, it should be measured by testing whether voters issue ballots for individual candidates or their party labels -- with each one of these measures painting an opposite picture of the relative strength of party ties.

One thing is clear: party ties at the electoral level are durable, with most voters remaining affiliated to a single party for their entire life, as highlighted by figure \emph{FIGURE}. Noting that party affiliations are registered at the municipal level, this empirical fact aligns with qualitative evidence provided by \citet{palmeira1995comicios}, who notes that parties at the local level constitute \emph{grupos pol\'{i}ticos} (political factions) with well-defined boundaries and power disputes. This relative stability of party ties at the electorate level for the minority of voters who are registered with a parties suggests a distinct dynamic tying an elite group of party members to the city hall.

\subsection{Municipal employment and patronage in Brazil}
